### 2011-08-23 paper.tex
first commit

% paper.tex - publication for Monthly Notices of the Royal Astronomical Society

\documentclass[useAMS,usenatbib]{mn2e}
\usepackage{txfonts}
\usepackage{graphicx}
\usepackage{natbib}

\def\del#1{{}}
%\def\del#1{{\bf (DELETED TEXT)}}
\def\C#1{#1}
%\def\C#1{{\bf #1}}

\sloppy

% --- macros --- %
\newcommand{\icm}{intra-cluster medium}
\newcommand{\cmb}{cosmic microwave background}

% --- journals --- %
\newcommand{\aap}{A{\&}A}
\newcommand{\aaps}{A{\&}AS}
\newcommand{\aj}{AJ}
\newcommand{\apj}{ApJ}
\newcommand{\apjl}{ApJL}
\newcommand{\apjs}{ApJS}
\newcommand{\apss}{Ap{\&}SS}
\newcommand{\araa}{ARA{\&}A}
\newcommand{\prd}{Phys. Rev. D}
\newcommand{\pre}{PRE}
\newcommand{\mnras}{MNRAS}
\newcommand{\ssr}{SSRv}
\newcommand{\nat}{Nature}
\newcommand{\physrep}{Phys.~Rep.}
\newcommand{\jcomp}{J.~Comput.~Phys.}

% --- christoph's commands --- %
\newcommand{\dd}{\mathrm{d}}
\newcommand{\bra}{\langle}
\newcommand{\ket}{\rangle}
\newcommand{\ltsima}{$\; \buildrel < \over \sim \;$}
\newcommand{\lsim}{\lower.5ex\hbox{\ltsima}}
\newcommand{\gtsima}{$\; \buildrel > \over \sim \;$}
\newcommand{\gsim}{\lower.5ex\hbox{\gtsima}}
\newcommand{\eps}{\varepsilon}

% --- Anders's commands --- %
\newcommand{\ev}{\mathrm{eV}}
\newcommand{\gev}{\mathrm{GeV}}
\newcommand{\mev}{\mathrm{MeV}}
\newcommand{\tev}{\mathrm{TeV}}
\newcommand{\mpc}{m_{\mathrm{p}}c}
\newcommand{\mpcc}{m_{\mathrm{p}}c^2}
\newcommand{\p}{\mathrm{p}}
\newcommand{\pp}{\mathrm{pp}}
\newcommand{\br}{\bold{r}}
\newcommand{\e}{\mathrm{e}}
\newcommand{\ii}{\mathrm{i}}
\newcommand{\inj}{\mathrm{inj}}
\newcommand{\pd}{\partial}
\newcommand{\CR}{\mathrm{CR}}
\newcommand{\gas}{\mathrm{gas}}
\newcommand{\mr}{\mathrm}
\newcommand{\DM}{\mathrm{DM}}
\newcommand{\ISM}{\mathrm{ISM}}
\newcommand{\ICM}{\mathrm{ICM}}
\newcommand{\CMB}{\mathrm{CMB}}
\newcommand{\IC}{\mathrm{IC}}
\newcommand{\mecc}{m_\e c^2}
\newcommand{\mug}{\mathrm{\mu G}}
\newcommand{\rvir}{R_{200}}
\newcommand{\B}{\mathrm{B}}
\newcommand{\ma}{\mathrm{max}}
\newcommand{\low}{\mathrm{low}}

% --- title --- %
\title{The origin of cosmic-ray electrons in cluster outskirts}

\author[A. Pinzke, P. Oh and C. Pfrommer] 
  {Anders Pinzke$^1$\thanks{e-mail:apinzke@physics.ucsb.edu (AP); peng@physics.ucsb.edu (PO);pfrommer@h-its.org (CP)}, Peng Oh$^1$\footnotemark[1],
    and Christoph Pfrommer$^2$\footnotemark[1]\\
    $^1$University of California - Santa Barbara,
  Department of Physics, CA 93106-9530, USA\\
    $^2$Heidelberg Institute for Theoretical Studies
  (HITS), Schloss-Wolfsbrunnenweg 33, DE - 69118 Heidelberg, Germany}

% --- document --- %
\begin{document}
\pagerange{\pageref{firstpage}--\pageref{lastpage}} \pubyear{2011}
\maketitle
\label{firstpage}

% --- abstract --- %
\begin{abstract}
  bla bla bla
\end{abstract}

% --- keywords --- %
\begin{keywords}
  magnetic fields, cosmic rays, radiation mechanisms: non-thermal, elementary
  particles, galaxies: cluster: general
\end{keywords}

% --- section: introduction --- %
\section{Introduction}

\section{Simulations}
\label{sect:simulations}
Our simulations were performed in a $\Lambda$CDM universe using the
cosmological parameters: $\Omega_{\mr m}=\Omega_\DM + \Omega_{\mr
  b}=0.3, \, \Omega_{\mr b}=0.039, \, \Omega_\Lambda=0.7, \, h=0.7, \,
n_{\mr s}=1$, and $\sigma_8=0.9$. The total matter density in units of
the critical density of the universe $\rho_{\rm crit}$ is denoted by
$\Omega_{\mr m}$, the baryonic density by $\Omega_{\mr b}$, the DM
density by $\Omega_\DM$ and the cosmological constant today is denoted
by $\Omega_\Lambda$. The critical density, $\rho_{\rm crit}=3H_0/(8\pi
G)$, where the present day Hubble constant $H_0=100\, h \, \mr{km}\,
\mr{s}^{-1} \, \mr{Myr}^{-1}$. $n_{\mr s}$ represents the spectral
index of the primordial power-spectrum, and $\sigma_8$ denotes the
$rms$ linear mass fluctuation within a sphere of radius $8 \,h^{-1}
\mr{Mpc}$ extrapolated to $z = 0$. The simulations were carried out
with an updated and extended version of the distributed-memory
parallel TreeSPH code GADGET-2 \citep{2005MNRAS.364.1105S,
  2001NewA....6...79S}. Gravitational forces were computed using a
combination of particle-mesh and tree algorithms.  Hydrodynamic forces
are computed with a variant of the smoothed particle hydrodynamics
(SPH) algorithm that conserves energy and entropy where appropriate,
i.e. outside of shocked regions \citep{2002MNRAS.333..649S}.  Our
simulations follow the radiative cooling of the gas, star formation,
supernova feedback, and a photo-ionizing background \citep[details can
  be found in][]{2007MNRAS.378..385P}. We model the cosmic ray (CR)
physics in a self-consistent way \citep{2006MNRAS.367..113P,
  2007A&A...473...41E, 2008A&A...481...33J}.  We include the adiabatic
CR transport process such as compression and rarefaction, and a number
of physical source and sink terms which modify the cosmic ray pressure
of each CR population separately. The most important sources
considered for injection are, diffusive shock acceleration at
cosmological structure formation shocks and shock waves in supernova
remnants, while the primary sinks are thermalization by Coulomb
interactions, and catastrophic losses by hadronization. For
simplicity, in this paper we do not take into account CRs injected
into the inter-stellar medium from supernova remnants (see Pinzke, Oh,
and Pfrommer, in prep. for a discussion of this topic).

\subsection{CR protons}
The CR proton distribution function is given by
\begin{equation}
  \label{eq:f_p}
  f_\p(p_\p) = \frac{\dd^2 N_\p}{\dd p_\p\,\dd V} = 
  C\, p_\p^{-\alpha}\,\theta(p_\p-q)\,, 
\end{equation}
and $p_\p = P_\p/m_\p c$, where we have normalized the proton momentum
$P_\p$ with the proton mass $m_\p$.  Here $q$ is the normalized
momentum cutoff, $\alpha$ the spectral index, and $C$ the
normalization in units of density.

\del{From the galaxy cluster simulations
we determine the adiabatic invariant Lagrangian amplitude of the
spectrum $\tilde{C}_0$, where the adiabatic changes of the gas density
$\rho$ change the normalization according to
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
and the convenient unitless redefinition of the amplitude is given by 
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}}

From our simulated galaxy clusters we derive for each snapshot at time
$t$ the injected CR proton distribution function $f_\p$. We select
particles in the outskirts of clusters where the gas densities are low
and hence the Coulomb cooling of the CR protons small. The CR proton
distribution function from the simulations is parameterized in terms of
adiabatic invariant momentum cutoff $q_0$ and the adiabatic invariant
Lagrangian amplitude of the spectrum $\tilde{C}_0$. We convert the adiabatic invariant quantities to physical quantities through
\begin{equation}
  \label{eq:q0}
  q=\left(\frac{\rho}{\rho_0}\right)^{\frac{1}{3}}q_0\,,
\end{equation}
and
\begin{equation}
  \label{eq:C0}
  \tilde{C}=\left(\frac{\rho}{\rho_0}\right)^{-\frac{\alpha-1}{3}}\tilde{C}_0\,,
\end{equation}
where the convenient unitless redefinition of the CR proton amplitude
is given by
\begin{equation}
  \label{eq:Ct}
  \tilde{C}=C\,m_\p/\rho\,.
\end{equation}
The injected distribution function is calculated as the change in
normalization between each snapshot, $f_\rmn{inj,p}(p_\p) = \Delta
C\,p_\p^{-\alpha}$, and is derived through the formalism developed
in \cite{2007A&A...473...41E} and \cite{2008A&A...481...33J}:
\begin{eqnarray}
  \label{eq:deltaC}
  \Delta C(t+\Delta t) &=& C(t) 
  \frac{\Delta \eps_\CR(t)-T_\p(q(t))\,\Delta n_\CR(t)}
       {\eps_\CR(t) - T_\p(q(t))\,n_\CR(t)}\,,\quad\rmn{where}\nonumber\\
       \Delta X(t) &=&  X(t+\Delta t) -  X(t)\,. 
\end{eqnarray}
We fix the time between each snapshot $\Delta t$ to $100$~Myrs, which
is smaller then the loss timescale of CR electrons in cluster
outskirts. The CR proton number density is given by
\begin{equation}
  \label{eq:ncr}
  n_{\CR} = \int_0^\infty \dd p_\p\, f_\p(p_\p) =
  \frac{C\, q^{1-\alpha}}{\alpha-1}\,,
\end{equation}
provided $\alpha >1$. The kinetic energy density of the CR
proton population is
\begin{eqnarray}
  \label{eq:epscr}
  \eps_\CR &=& \int_0^\infty  \dd p_\p\, f_\p(p_\p) \,T_{\rmn
    p}(p_\p)=\frac{C\, m_\p\,c^2}{\alpha-1} \, \times
  \nonumber \\
  && \left[\frac{1}{2}
    \, \B_{\frac{1}{1+q^2}} \left(
    \frac{\alpha-2}{2},\frac{3-\alpha}{2}\right) + q^{1-\alpha}
    \left(\sqrt{1+q^2}-1 \right) \right] \,,
\end{eqnarray}
where $T_\p(p_\p) = (\sqrt{1+p_\p^2} -1)\, m_\p\,c^2$ is the kinetic energy
of a CR proton. $\B_x(a,b)$ denotes the incomplete Beta-function, and
$\alpha>2$ is assumed.

\section{Radio relics}
\label{section:relics}
RELICS...  Collisionless cluster shocks are able to accelerate ions
and electrons in the high-energy tail of their Maxwellian distribution
functions through diffusive shock acceleration \citep[for reviews
  see][]{1983RPPh...46..973D, 1987PhR...154....1B,
  2001RPPh...64..429M}. Neglecting non-linear shock acceleration and
cosmic ray modified shock structure, then electrons and protons are
indistinguishable in the process of diffusive shock
acceleration. There are, however, difference in: (1) the maximum
energy of the steady state spectrum that depends on the details of the
shock, (2) acceleration efficiency that is related to the smaller
Larmor radius of the electrons that keeps the electrons from diffusing
back and forth over the discontinuity of the shock front.CHECK

In this section we derive the CR electron distribution function by
rescaling the injected CR proton spectrum to account for the different
acceleration efficiencies as well as the shift in momentum due to the
factor $\sim 2000$ difference in mass. In addition we model the
Coulomb and radiative losses of the CR electrons.

\subsection{CR electron distribution function} 
The injected CR electron distribution function at each time $t$ is
given by
\begin{eqnarray}
  \label{eq:f_inj_e}
  f_\rmn{inj,e}(p_\e) &=&  \frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd p_\e}(p_\e) = 
\frac{\dd^2 N_\rmn{inj,e}}{\dd V \dd p_\p}(p_\e)\frac{m_\e}{m_\p} \nonumber\\
&=& \frac{\dd^2 N_\rmn{inj,p}}{\dd V \dd p_\p}(p_\p)
\left(\frac{p_\e}{p_\p}\right)^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}\frac{m_\e}{m_\p} = 
\Delta C\,p_\e^{-\alpha}\frac{\eta_\rmn{max,e}}{\eta_\rmn{max,p}}\frac{m_\e}{m_\p}\,.
 \nonumber\\
&&
\end{eqnarray}
and $p_\e = P_\e/m_\e c$, where we have normalized the electron
momentum $P_\e$ with the electron mass $m_\p$. We use that maximal
50\% of the energy available in a shock is injected into CR protons
($\eta_\rmn{max,p}$) and a factor 10 smaller efficiency of 5\% for the
CR electrons ($\eta_\rmn{max,e}$).

The CR electrons cool through inverse Compton (IC) emission and
Coulomb losses on timescales that are relative short compared to the
dynamical timescale of a cluster. We model these looses analytically
by assuming a self similar solution for the CR electron distribution
function \citep{1999ApJ...520..529S}
\begin{eqnarray}
  \label{eq:f_inj_e}
  f_\e(p_\e) &\approx& \frac{f_\rmn{inj,e}(\gamma_\e)}{\alpha-1}
\left(\frac{\gamma_\low}{\gamma_\e}+\frac{\gamma_\e }{\gamma_\ma}\right)^{-1}
\nonumber\\
&\times& \left\{2-\left[\frac{1}{2}\left(1-\frac{\gamma_\e}{\gamma_\ma}+
\left|1-\frac{\gamma_\e}{\gamma_\ma}\right|\right)\right]^{\alpha-1}\right.
\nonumber\\
&-& \left.\left(1+\frac{\gamma_\low}{\gamma_\e}\right)^{-(\alpha-1)}\right\}\,,
\end{eqnarray}
Here $\gamma_\e \approx p_\e = P_\e/m_\e c$ for $ p_\e \gg 1$.  The
approximation in Eqn.~(\ref{eq:f_inj_e}) is valid as long as
$\gamma_\ma\gg \gamma_\low$, where the upper limit is determined from
IC losses through
\begin{eqnarray}
  \label{eq:gamma_max}
b_\IC(\gamma_\e) &=& b_\IC \gamma_\e^2 
= \frac{4}{3}\frac{\sigma_\rmn{T}}{m_\e c} U_\rmn{cmb} \gamma_\e^2 \,,\quad \rmn{and}\nonumber \\
%1.37\times 10^{-20} \gamma^2(1+z)^4 \,\rmn{s}^{-1}
\gamma_\ma &=& \frac{1}{b_\IC\,\Delta t}\,.
\end{eqnarray}
Here $\sigma_\rmn{T}= 8\pi e^4/3(m_\e c^2)^2$ is the Thomson cross
section and $U_\rmn{cmb}$ is the energy density of the CMB in the
cluster. The lower limit is determined from Coulomb losses through
\begin{eqnarray}
  \label{eq:gamma_min}
b_\rmn{C}(\gamma_\e) &=& b_\rmn{C} \gamma_\e^2 
= \frac{3\,\sigma_\rmn{T}\,n_\rmn{el}\,c}{2}
\left[\ln\left(\frac{m_\e c^2 \sqrt{\gamma_\e-1}}{\hbar\,\omega_\rmn{plasma}}\right)\right.\nonumber \\
  &-&\ln(2)\left(\frac{1}{2}+\frac{1}{\gamma_\e}\right)+\frac{1}{2}+
\left.\left(\frac{\gamma_\e-1}{4\gamma_\e}\right)^2\right]
         \,,\quad \rmn{and}\nonumber \\
\gamma_\low &=& b_\rmn{C}\,\Delta t\,.
\end{eqnarray}
Here $\omega_\rmn{plasma} = \sqrt{4\pi e^2 n_\e / m_\e}$ is the plasma
frequency, and $n_\e$ is the number density of free electrons.

We model the final CR electron spectrum as a superposition of five CR
populations, each determined by Eqn.~\ref{eq:f_inj_e} but with a
different spectral index and injected spectra.

\subsection{Synchrotron emission}
In this section we derive the synchrotron emission from the CR
electron distribution.  WORK UNDER PROGRESS

\del{Collisionless cluster shocks are able to accelerate ions and electrons
in the high-energy tail of their Maxwellian distribution functions
through diffusive shock acceleration \citep[for reviews
  see][]{1983RPPh...46..973D, 1987PhR...154....1B,
  2001RPPh...64..429M}.  Neglecting non-linear shock acceleration and
cosmic ray modified shock structure, the process of diffusive shock
acceleration uniquely determines the spectrum of the freshly injected
relativistic electron population in the post-shock region that cools
and finally diminishes as a result of loss processes. \del{The inverse
Compton emitting electron population cools on such a short time scale
$\tau_\rmn{sync} < 10^8$~yrs (compared to the very long dynamical time
scale $\tau_\rmn{dyn} \sim 1$~Gyr) that we can describe this by
instantaneous cooling. In this approximation, there is no steady-state
electron population and we would have to convert the energy from the
electrons to inverse Compton (IC) and synchrotron radiation. Instead,
we introduce a virtual electron population that lives in the
SPH-broadened shock volume only; this is defined to be the volume
where energy dissipation takes place. Within this volume, which is
co-moving with the shock, we can use the steady-state solution for the
distribution function of relativistic electrons and we assume no
relativistic electrons in the post-shock volume, where no energy
dissipation occurs. Thus, the cooled CR electron equilibrium spectrum
can be derived from balancing the shock injection with the
IC/synchrotron cooling: above a GeV it is given by}
\begin{equation}
\label{eq:f_eq}
f_\e(E) = C_\e\, E^{-\alpha_\e}, \quad
C_\e \propto \frac{\rho}
{\eps_B + \eps_\rmn{ph}}
\end{equation}
}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/CRespec0.3Rv.140.v12.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/CRespec1.4Rv.140.v12.eps}
  \caption{Cosmic ray electron spectra in cluster outskirts at
    redshift zero. We show the CR electron distribution function
    weighted with the electron Lorentz factor ($\gamma_\e$)
    squared. In the {\it left panel} panel we show the CR electron in
    the region between (0.3-0.5)$\rvir$ and in the {\it right panel}
    between (1.3-1.5)$\rvir$. The solid lines show the median CR
    electron spectrum where both radiative and adiabatic losses are
    included. The dotted lines shows the 68 percentiles. The dashed
    lines show the CR electron spectrum without Coulomb and radiative
    cooling. Note that the Coulomb cooling at low energies is very
    inefficient because of the low electron densities in the cluster
    outskirts. \label{fig:e_spec}}
\end{minipage}
\end{figure*}

high energy part build up during last Gyr, also mention normalization
\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/f_evolution.0.3Rv.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/f_evolution.1.4Rv.eps}
  \caption{Time evolution of cosmic ray (CR) electron spectra. We show
    the CR electron distribution function weighted with the electron
    Lorentz factor ($\gamma_\e$) squared. The time evolution of the CR
    electrons which at $z=0$ reside from (0.3-0.5)$\rvir$ are shown in
    the {\it left panel} and (1.3-1.5)$\rvir$ in the {\it right
      panel}. The black lines show the CR electrons with a
    $\gamma_\e=30$, blue lines $\gamma_\e=200$, and green lines
    $\gamma_\e=10^4$. The solid lines show the CR electrons with full
    cooling, i.e. for Coulomb, inverse Compton, and adiabatic
    losses. Dashed lines show the CR electrons without Coulomb and
    inverse Compton cooling while the dash-dotted lines show the CR
    electrons without adiabatic cooling. The dotted line show the
    injected CR electron distribution function, derived from the
    injected CR protons. For each energy, we normalize the CR electron
    distribution function with the CR electrons without
    cooling. Notice that the high energy part is built up during the
    last Gyr. This is explained by the inverse Compton losses that are
    larger at high energies, especially at high redshifts where the
    energy density of the CMB is much larger than
    today.  \label{fig:e_evol}}
\end{minipage}
\end{figure*}

\begin{figure*}
\begin{minipage}{2.0\columnwidth}
  \includegraphics[width=0.49\columnwidth]{./figures/evolution.0.3Rv.eps}
  \includegraphics[width=0.49\columnwidth]{./figures/evolution.1.4Rv.eps}
  \caption{Time evolution of the gas density and the cosmic ray (CR)
    cooling time. We show the time evolution of the electron number
    density and CR electron cooling time that are associated with the
    CR electrons which at redshift $z=0$ reside from (0.3-0.5)$\rvir$
    in the {\it left panel} and (1.3-1.5)$\rvir$ in the {\it right
      panel}. The red lines show the electron number densities, where
    the dashed line show the median and dash-dotted the mean. The
    solid lines show the cooling times
    $\tau_\rmn{cool}=\gamma_\e/[b_\IC(\gamma_\e)+b_\rmn{C}(\gamma_\e)]$
    for $\gamma_\e=30$ (black), $\gamma_\e=200$ (blue), and
    $\gamma_\e=10^4$ (green). Interesting the CR cooling at low
    energies of the particles that end up in the center is constant
    while is increasing monotonically for the particles that end up in
    the cluster periphery. The density of the particles that end up in
    the center is roughly constant with time while it is decreasing
    for the particles that end up in the outer part. This is due to
    the expansion of the universe that counteract the contraction of
    the particles in the cluster periphery. \label{fig:evol}}
\end{minipage}
\end{figure*}

% --- section: --- %
\section{Conclusions}
\label{sect:conclusions}

% --- section: --- %
\section*{Acknowledgments}

% --- section: bibliography --- %
\bibliography{bibtex/paper}
\bibliographystyle{mn2e}

\appendix

\bsp

\label{lastpage}

\end{document}